---
title: "Comparison of life expenctancy for three countries"
author: "your name"
params:
  country_1: Belgium
  country_2: !r sample(c("United States", "Mexico", "Canada"), 1)
---


```{r load-packages}
library(tidyverse)
```

```{r load-data}
gap <- read_csv("data/gapminder.csv")
```

Define a function, `plot_country`, that takes country names as one of its arguments and outputs a plot of life expectancy by year for each country and displays the GDP per capita of the country as size of the points.

```{r def-plot-country}
plot_country <- function(gap, ctry) {
  ctry <- match.arg(ctry, unique(gap$country),  several.ok = TRUE)
  gap %>%
    filter(country %in% ctry) %>%
    ggplot(aes(x = year, y = lifeExp, colour = country, size = gdpPercap)) +
    geom_point() +
    labs(title = paste(ctry, collapse = ", "),
         subtitle = "Change in life expectancy and GDP per capita")
}

country_3 <- "South Africa"
```

We compare the change in life expectancy for `r params$country_1`, `r params$country_2` and `r country_3`:

```{r plot-country, echo=FALSE}
plot_country(gap, c(params$country_1, params$country_2, country_3))
```

```{r table-country}
gap %>%
  filter(country %in% c(params$country_1, params$country_2, country_3)) %>%
  select(country, lifeExp, pop,  gdpPercap) %>%
  group_by(country) %>%
  summarise_all(mean) %>%
  knitr::kable()
```

A few new things are happening in this document:

- The use of parameters to change things in only 1 place
- The use of R functions within parameters with the `!r` notation
- The use of functions to generate plots, making it easier to do:
    - type checking
    - reduce code repetition
- The use of `knitr::kable` to generate nicely formatted (simple) tables within Rmarkdown documents

